\section{Hypothesis}
\label{sec:hypothesis}

\subsection{Hypothesis}
The main focus of this research is to see if equal accuracy can be achieved with a pre-processed representation of the original community graph using a combination of previously existing algorithms.
These algorithms have further been explained in Section \ref{sec:relevantResearch}.

\subsection{Goal}
The goal of this research is to see if we can achieve similar performance with a reduced version of the original graph. 
Our expectation is that our pre-processing algorithm will significantly reduce the size of the graph if it is highly interconnected. 
We expect to see graphs with a higher ratio of edges to vertexes to be more reduced. 
In turn, we expect the genetic algorithm to be faster when processing a reduced graph compared to the original graph.

\subsection{Research questions}
\subsection*{Main research question}
\begin{quote}
Does the reduction of a graph have any influences on the performance of current state of the art, mainly on the time that is required to reach reasonably good results?
\end{quote}
\subsection*{Other research questions}
To reduce the complexity of the main research question, some smaller research questions have been chosen.
\begin{quote}
What kind of influence does the replacement of a small cliques by a single node have on the information in the graph?

Can all of the information of the graph be retained while still reducing the size?

Is it worth to pre-process a graph before applying the meta-heuristic?

What differences are there between the algorithm with and without the pre-processing?
\end{quote}
